PhD Title \\ PhD Thesis

### 2024-07-04 Latex_files/Thesis/chapters/chapter 1/main.tex
updates

>>>>

\subsection{Shear induced Nucleation}
It has long been known that solution nucleation is influenced by shearing 
of the fluid, from stirring agents in the container (i.e. propellers) to 
the container boundary layers, however there has yet to be a clear mechanism 
by which said shearing effects the nucleation event. Theoretical research 
into shear induced nucleation suggests that there should be a slight increase 
in the nucleation rate at low shear rates, reaching a maximum increase in 
nucleation rate, and then at higher shear rates the nucleation rate begins 
to drop off. This has been shown theoretically for both simple colloidal
\cite{Mura2016,Debuysschere2023,Richard2015} and ice crystal formation 
\cite{Goswami2020}; however, no experimental work into these systems has 
been conducted to prove this is the case.  There is some experimental 
evidence for this phenomena in simple salt and protein solutions - though 
the authors emphasise that mechanical agitation cannot be ruled out - 
there has not been a exhaustive study into the shearing effects apart from 
in glycine solutions. In \cite{Debuysschere2023} it was found that a shear 
rate of around $3000\ s^{-1}$ was the maximum shear rate that would yield 
the highest nucleation rate. Using the theoretical model established in 
\cite{Mura2016,2001} which modifies the CNT to account for the effects of 
a nucleus undergoing shearing, accounting for the fact that a nucleus' 
growth is undergoing competition between flow-mediated molecular transport 
and the strain applied by the flow field which inhibits the growth of the 
nucleus. There central conclusion (from both the theoretical and experimental 
results) is that there is an optimal shear rate in which the nucleation 
rate is maximised. However, a question that arises from this result, if 
there is a optimal shear rate in which molecular transport is maximised 
and strain is minimised, then surely there should also be a shear rate 
in which the molecular transport and strain are equal - allowing one to 
suspend a nucleus at a constant radius. In this scenario, the molecular 
transport would prevent the nucleus from dissolving, but the strain would 
prevent the nucleus from growing. This however would require one to be 
able to apply a continuous shear rate to a targeted nucleus with high 
precision, there is also no model for an individual nucleus undergoing growth.

\section{Aims and Objectives}
Overall the aim of the PhD is to study viability of using micro-rotors to
generate localised fluid flow around the beam focus. The results are reported
in chapter 3, this is then succeeded by experimental work where we 
instead using a galvano-mirror to generate a moving beam focus within an 
undersaturated solution to control the crystal growth front. In addition,
computer simulations into asymmetric dimers and their applications as complex
micro-rotors are discussed in chapter 4. Lastly two novel methods of measuring
rotational motion are discussed in chapter 5; firstly via a novel detection
fibre method that allows for instantaneous measurements of the orientational
behaviour of optically trapped ellipsoids/dimers; and secondly we create 
a simulative quadrant photo diode that replicates laboratory results, utilising 
linear regression techniques we measure the change in orientation in order to
measure the optical torque applied to a non-birefringent particle.

### 2024-07-04 Latex_files/Thesis/chapters/chapter 3/main.tex
updates

Vaterite is a polymorph of calcium carbonate that is rarely seen in 
nature due to its low stability. However unlike its other polymorphs of 
calcite and aragonite, when synthesised vaterite will typically form 
small spherical particles making them ideal for optical trapping. 
Synthesis of vaterite micro spheres requires fine control of the nucleation
process in order to maintain polymorphic control.

@@

\begin{figure}
	\centering
	\begin{subfigure}{0.4\linewidth}
		\includegraphics[width=\linewidth]{vaterite_sample.jpg}
		\subcaption{}
	\end{subfigure}
	\begin{subfigure}{0.55\linewidth}
		\includegraphics[width=\linewidth]{rotating_psd.png}
		\subcaption{}
	\end{subfigure}
	\caption{(a)Sample vaterite sphere suspended in water and trapped by circular polarised trap. (b) Collected power spectrum from rotating vaterite, peaks in the power spectrum appear at integer multiples of the rotational frequency ($f_{rot} \approx 49.8$ Hz)}
\end{figure}

\subsection{Rotation of Vaterite micro spheres}

\section{Shear induced Nucleation}
\comment{This belongs in the introduction chapter me thinks}

It has long been known that solution nucleation is influenced by shearing of the fluid, from stirring agents in the container (i.e. propellers) to the container boundary layers, however there has yet to be a clear mechanism by which said shearing effects the nucleation event. Theoretical research into shear induced nucleation suggests that there should be a slight increase in the nucleation rate at low shear rates, reaching a maximum increase in nucleation rate, and then at higher shear rates the nucleation rate begins to drop off. This has been shown theoretically for both simple colloidal \cite{Mura2016,Debuysschere2023,Richard2015}  and ice crystal formation \cite{Goswami2020}; however, no experimental work into these systems has been conducted to prove this is the case.  There is some experimental evidence for this phenomena in simple salt and protein solutions - though the authors emphasise that mechanical agitation cannot be ruled out - there has not been a exhaustive study into the shearing effects apart from in glycine solutions. In \cite{Debuysschere2023} it was found that a shear rate of around $3000\ s^{-1}$ was the maximum shear rate that would yield the highest nucleation rate. Using the theoretical model established in \cite{Mura2016,2001} which modifies the CNT to account for the effects of a nucleus undergoing shearing, accounting for the fact that a nucleus' growth is undergoing competition between flow-mediated molecular transport and the strain applied by the flow field which inhibits the growth of the nucleus. There central conclusion (from both the theoretical and experimental results) is that there is an optimal shear rate in which the nucleation rate is maximised. However, a question that arises from this result, if there is a optimal shear rate in which molecular transport is maximised and strain is minimised, then surely there should also be a shear rate in which the molecular transport and strain are equal - allowing one to suspend a nucleus at a constant radius. In this scenario, the molecular transport would prevent the nucleus from dissolving, but the strain would prevent the nucleus from growing. This however would require one to be able to apply a continuous shear rate to a targeted nucleus with high precision, there is also no model for an individual nucleus undergoing growth.

Optical tweezing has often been used for micro-rheology, by computing the
exact forces being exerted on the trapped sphere, one can determine the
local temperature/viscosity of the medium \cite{Millen2014, RodriguezSevilla2018}.
Typically one would use a birefringent particle (i.e. vaterite, liquid
crystals, etc) and rotating it within the fluid, the maximum rotation
rate being a product of the fluid drag resisting the torque of the 
trapping beam \cite{RodriguezSevilla2018}. Or if you want to measure 
fluid flow you can instead use a micro-rotor to see how fluid flow 
propagates in the medium \cite{Knoener2005}. Likewise, one can use 
a galvanometric mirror to probe the drag force of the fluid, by 
understanding the trap strength (calibrating using a low frequency 
signal) one can measure the drag force experienced by the local fluid \cite{RobertsonAnderson2018}. I

@@

First we tested the rotational behaviour of vaterite in distilled water, 
samples of vaterite where diluted down and $200\mu L$ was pipetted onto
the sample stage. Due to Van der Waal's forces some of the microspheres 
were stuck together, fortunately individual sphere's were still

@@

Vaterite samples were synthesised according to \cite{Parkin2009, Bishop2004} 
(see sec.~\ref{sec:vaterite}), and then suspended in a droplet of supersaturated solutions of Glycine and water. After locating a singular microsphere the
vaterite was trapped in a circularly polarised light and brought close
to the droplet edge, after a period of ten minutes if no nucleation event
was observed the particle was released and another particle was trapped. 
Due to the increased viscosity of the supersaturated solution and the 
proximity to the droplet edge, the observed rotation rate 
was rather low or non-existent, the results are tabulated below:

\begin{table}
	\centering
	\caption{Results from rotating vaterite within supersaturated solution. Solubility concentration for Glycine was at $16^\circ$, $C^*=0.2016g/g$}
	\begin{tabular}[width=\textwidth]{|c|c|c|c|}
		\hline
		Super Saturation & Particle radius [$\mu$] & $\omega$ [Hz] & Nucleation [$\checkmark/\times$]\\
		\hline
		\multirow{3}*{1.01} & 2.34 & 20.7 & $\times$ \\
		\cline{2-4} & 5.67 & 23.3 & $\times$ \\
		\cline{2-4} & 3.26 & 15.4 & $\times$ \\
		\hline
		\multirow{3}*{1.14} & 1.89 & 1.23 & $\times$ \\
		\cline{2-4} & 3.75 & 3.54 & $\times$ \\
		\cline{2-4} & 4.35 & 4.86 & $\times$ \\
		\hline
		\multirow{3}*{1.29} & 3.47 & 0.00 & $\times$ \\
		\cline{2-4} & 1.59 & 0.00 & $\times$ \\
		\cline{2-4} & 6.24 & 0.00 & $\times$ \\
		\hline
		\multirow{3}*{1.40} & 6.32 & 0.00 & $\times$ \\
		\cline{2-4} & 3.68 & 0.00 & $\times$ \\
		\cline{2-4} & 5.43 & 0.00 & $\times$ \\
		\hline
		\multirow{3}*{1.49} & 4.76 & $0.00$ & $\times$ \\
		\cline{2-4} & 7.27 & $0.00$ & $\times$ \\
		\cline{2-4} & 1.52 & $0.00$ & $\times$ \\
		\hline
	\end{tabular}
\end{table}

This would explain partly why we could never seen nucleation even in the case where our vaterite could rotate freely. The closest we could trap a microsphere to the droplet edge was in the range of $5-10\mu\ m$, at that distance the fluid flow would be so low that not even using a liquid droplet rotor would achieve the rotational speeds necessary to localise nucleation.

>>>>

Vaterite is a polymorph of calcium carbonate that is rarely seen in 
nature due to its low stability. However unlike its other polymorphs of 
calcite and aragonite, when synthesised vaterite will typically form 
small spherical particles making them ideal for optical trapping and rotation. 
Synthesis of vaterite micro spheres requires fine control of the nucleation
process in order to maintain polymorphic stability, though for the purposes
of optical rotation the exact polymorph is not as important as its morphology
as all 3 polymorphs are inherently birefringent.

@@

\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.4\linewidth}
		\includegraphics[width=\linewidth]{vaterite_sample.jpg}
		\subcaption{}
	\end{subfigure}
	\begin{subfigure}{0.55\linewidth}
		\includegraphics[width=\linewidth]{rotating_psd.png}
		\subcaption{}
	\end{subfigure}
	\caption{(a)Sample vaterite sphere suspended in water and trapped by circular polarised trap. (b) Collected power spectrum from rotating vaterite, peaks in the power spectrum appear at integer multiples of the rotational frequency ($f_{rot} \approx 49.8$ Hz)}
\end{figure}

Liquid crystals are an intriguing example of materials with mixed phase
properties, unlike typical solutes such as Glycine, a liquid crystal
can still maintain some degree of order between its individual molecules
while in the liquid state. This is due to the fact that liquid crystals
are constructed of long ordered molecules that demonstrate a long range
ordering. There are three main types of liquid crystal transition methods:
thermotropic crystals will transition to their liquid crystal phase when 
sufficiently heated; Lyotropic materials can undergo this transition due 
to changes in temperature and concentration; and lastly Metallotropic 
materials - which are composed of both organic and inorganic molecules - 
change phase according to the ratio of organic to inorganic molecules present.
Liquid crystal rotors are rather simple in their production, 
4-Heptyl-4-biphenylcarbonitrile (7CB) was purchased from Sigma Aldrich
and a small amount was added to a vial of distilled water. The solution 
was then heated in a water bath to $25^\circ$ in order to transition the
solid crystal into its liquid crystal state. The individual droplets 
are inherently birefringent and would rotate readily while suspended in 
distilled water.

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.5\linewidth]{LC_sample.png}
	\caption{Liquid crystal undergoing rotation due to the circularly polarised trap.} 
\end{figure}
\subsection{Rotation of Vaterite micro spheres} 
Optical tweezing has often been used for micro-rheology, by computing the
exact forces being exerted on the trapped sphere, one can determine the
local temperature/viscosity of the medium \cite{Millen2014, RodriguezSevilla2018}.
Typically one would use a birefringent particle (i.e. vaterite, liquid
crystals, etc) and rotating it within the fluid, the maximum rotation
rate being a product of the fluid drag resisting the torque of the 
trapping beam \cite{RodriguezSevilla2018}. Or if you want to measure 
fluid flow you can instead use a micro-rotor to see how fluid flow 
propagates in the medium \cite{Knoener2005}. Likewise, one can use 
a galvanometric mirror to probe the drag force of the fluid, by 
understanding the trap strength (calibrating using a low frequency 
signal) one can measure the drag force experienced by the local fluid \cite{RobertsonAnderson2018}. I

@@

First we tested the rotational behaviour of vaterite in distilled water, 
samples of vaterite where diluted down and $200\mu L$ was pipetted onto
the sample stage. Due to Van der Waal's forces some of the microspheres 
were stuck together, fortunately individual sphere's were still present.
Over the course of a few weeks multiple sphere's where trapped 
and their rotation speeds were measured via the QPD.

@@

\section{Rotation of Vaterite in supersaturated solutions}

Vaterite samples were synthesised according to \cite{Parkin2009, Bishop2004} 
(see sec.~\ref{sec:vaterite}), and then suspended in a droplet of 
supersaturated solutions of Glycine and water. After locating a 
singular microsphere the vaterite was trapped in a circularly 
polarised light and brought close to the droplet edge, after a 
period of ten minutes if no nucleation event was observed the 
particle was released and another particle was trapped. Due to 
the increased viscosity of the supersaturated solution and the 
proximity to the droplet edge, this was done as previous work with
optical tweezers have only seen nucleation near the solution-air interface
\cite{Gowayed2021, Liao2022, Walton2018}. The observed rotation rate 
was rather low or non-existent, the results are tabulated below:
\begin{table}
	\centering
	\caption{Results from rotating vaterite within supersaturated solution of $H_2O$ and Glycine. Solubility concentration for Glycine at $16^\circ$ was $C^*=0.2016g/g$}
	\begin{tabular}[width=\textwidth]{|c|c|c|c|}
		\hline
		Super Saturation & Particle radius [$\mu$] & $\omega$ [Hz] & Nucleation [$\checkmark/\times$]\\
		\hline
		\multirow{3}*{1.01} & 2.34 & 20.7 & $\times$ \\
		\cline{2-4} & 5.67 & 23.3 & $\times$ \\
		\cline{2-4} & 3.26 & 15.4 & $\times$ \\
		\hline
		\multirow{3}*{1.14} & 1.89 & 1.23 & $\times$ \\
		\cline{2-4} & 3.75 & 3.54 & $\times$ \\
		\cline{2-4} & 4.35 & 4.86 & $\times$ \\
		\hline
		\multirow{3}*{1.29} & 3.47 & 0.00 & $\times$ \\
		\cline{2-4} & 1.59 & 0.00 & $\times$ \\
		\cline{2-4} & 6.24 & 0.00 & $\times$ \\
		\hline
		\multirow{3}*{1.40} & 6.32 & 0.00 & $\times$ \\
		\cline{2-4} & 3.68 & 0.00 & $\times$ \\
		\cline{2-4} & 5.43 & 0.00 & $\times$ \\
		\hline
		\multirow{3}*{1.49} & 4.76 & $0.00$ & $\times$ \\
		\cline{2-4} & 7.27 & $0.00$ & $\times$ \\
		\cline{2-4} & 1.52 & $0.00$ & $\times$ \\
		\hline
	\end{tabular}
\end{table}

The calculations from fig.\ref{fig:vaterite_shear} indicate that even
in pure water the rotational speed achieved by a microsphere is
insufficient to achieve a significant shear rate more than a few microns
from its surface. The closest we could trap a microsphere to the droplet
edge was in the range of $5-10\mu\ m$, at that distance the fluid flow
would be so low that not even using a liquid droplet rotor would achieve
the rotational speeds necessary to localise nucleation. While in theory
a sufficiently focused laser could rotate any microsphere to a fast enough 
to match the speeds expected by \cite{Debuysschere2023} the localised intensity
would be so large that even using $D_2O$ would see a significant increase in 
temperature.

\section{Shearing via a Galvano-mirror}
